---
title: "DS6306 Case Study 2"
author: "Andre Mauldin"
date: "8/1/2020"
output: html_document
---
## In this study we are analyzing causes of attrition.  We will also set up a model to predict future attrition.
```{r}
library(magrittr)
library(ggplot2)
library(reshape2)
library(maps)
library(dplyr)
library(tidyr)
library(tidyverse)
library(usmap)
library(GGally)
library(ggthemes)
library(gridExtra)
library(knitr)
library(kableExtra)
library(stringr)
library(class)
library(caTools)
library(caret)
library(e1071)
library(plotly)
library(openxlsx)
turnoverData = read.csv('CaseStudy2-data.csv')

```
## Helper Functions 
These helper functions aid in aggregating the data to set up the tables and charts.
```{r}
turnover = data.frame(turnoverData)
turnover = turnover %>% filter(Attrition == "Yes")
turnover = data.frame(turnover)

turnoverFunc <- function(columnName, data, colString) {
  arguments <- as.list(match.call())
  columnName <- eval(arguments$columnName, data)
  df = turnover %>% 
    group_by(Attrition, as.factor(columnName)) %>%
    summarize(count = n())
  df = arrange(df, desc(count))
  # df = head(df, 10)
  df = data.frame(df)
  colnames(df) = c("Attrition", "AttritionValue", "AttritionCount")
  dfplot = df %>% ggplot(aes(x=AttritionValue, y=AttritionCount)) + geom_col() + xlab(colString)
  return(ggplotly(dfplot))
}

turnoverMax <- function(columnName, data, colString) {
  arguments <- as.list(match.call())
  columnName <- eval(arguments$columnName, data)
  df = turnover %>% 
    group_by(Attrition, as.factor(columnName)) %>%
    summarize(`Attrition Count` = n()) # %>%
    # filter(`Attrition Count` == max(`Attrition Count`))
  
  # df = arrange(df, desc(count))
  # df = head(df, 10)
  # df = data.frame(df)
  colnames(df) = c("Attrition", "Attrition Factor", "Attrition Count")
  df = df %>% mutate(Reason = colString)
  return(df)
}

```
## All Attrition Factors
This table shows all Reasons and factors contributing to attrition.  The Reasons are the column names from the data set.  The factors are the factors in those columns. The Attrition Count is the number of times Attrittion equaled "Yes" per Attrition Factor.  It is sorted descending by Attrition Factor.
```{r}
AgeColumn = turnoverMax(Age, data=turnover, "Age")
BusinessTravelColumn = turnoverMax(BusinessTravel, data=turnover, "Business Travel")

turnoverFactors = rbind(AgeColumn, BusinessTravelColumn)

DepartmentColumn = turnoverMax(Department, data=turnover, "Department")
turnoverFactors = rbind(turnoverFactors, DepartmentColumn)

DistanceFromHomeColumn = turnoverMax(DistanceFromHome, data=turnover, "Distance from Home")
turnoverFactors = rbind(turnoverFactors, DistanceFromHomeColumn)

EducationColumn = turnoverMax(Education, data=turnover, "Education")
turnoverFactors = rbind(turnoverFactors, EducationColumn)

EducationFieldColumn = turnoverMax(EducationField, data=turnover, "Education Field")
turnoverFactors = rbind(turnoverFactors, EducationFieldColumn)

EnvironmentSatisfactionColumn = turnoverMax(EnvironmentSatisfaction, data=turnover, "Emvironment Satisfaction")
turnoverFactors = rbind(turnoverFactors, EnvironmentSatisfactionColumn)

GenderColumn = turnoverMax(Gender, data=turnover, "Gender")
turnoverFactors = rbind(turnoverFactors, GenderColumn) 

HourlyRateColumn = turnoverMax(HourlyRate, data=turnover, "Hourly Rate")
turnoverFactors = rbind(turnoverFactors, HourlyRateColumn) 

JobInvolvementColumn = turnoverMax(JobInvolvement, data=turnover, "Job Involvment")
turnoverFactors = rbind(turnoverFactors, JobInvolvementColumn) 
  
JobLevelColumn = turnoverMax(JobLevel, data=turnover, "Job Level")
turnoverFactors = rbind(turnoverFactors, JobLevelColumn)
  

JobRoleColumn = turnoverMax(JobRole, data=turnover, "Job Role")
turnoverFactors = rbind(turnoverFactors, JobRoleColumn)

JobSatisfactionColumn = turnoverMax(JobSatisfaction, data=turnover, "Job Satisfaction")
turnoverFactors = rbind(turnoverFactors, JobSatisfactionColumn) 

MaritalStatusColumn = turnoverMax(MaritalStatus, data=turnover, "Marital Status")
turnoverFactors = rbind(turnoverFactors, MaritalStatusColumn)

MonthlyIncomeColumn = turnoverMax(MonthlyIncome, data=turnover, "Monthly Income")
turnoverFactors = rbind(turnoverFactors, MonthlyIncomeColumn)

MonthlyRateColumn = turnoverMax(MonthlyRate, data=turnover, "Monthly Rate")
turnoverFactors = rbind(turnoverFactors, MonthlyRateColumn)

NumCompaniesWorkedColumn = turnoverMax(NumCompaniesWorked, data=turnover, "Number of Companies Worked")
turnoverFactors = rbind(turnoverFactors, NumCompaniesWorkedColumn) 

Over18Column = turnoverMax(Over18, data=turnover, "Over 18")
turnoverFactors = rbind(turnoverFactors, Over18Column)

OverTimeColumn = turnoverMax(OverTime, data=turnover, "Overtime")
turnoverFactors = rbind(turnoverFactors, OverTimeColumn)

PercentSalaryHikeColumn = turnoverMax(PercentSalaryHike, data=turnover, "Percent Salary Hike")
turnoverFactors = rbind(turnoverFactors, PercentSalaryHikeColumn)  

PerformanceRatingColumn = turnoverMax(PerformanceRating, data=turnover, "Performance Rating")
turnoverFactors = rbind(turnoverFactors, PerformanceRatingColumn)

RelationshipSatisfactionColumn = turnoverMax(RelationshipSatisfaction, data=turnover, "Relationship Satisfaction")
turnoverFactors = rbind(turnoverFactors, RelationshipSatisfactionColumn)

StandardHoursColumn = turnoverMax(StandardHours, data=turnover, "Standard Hours")
turnoverFactors = rbind(turnoverFactors, StandardHoursColumn)

StockOptionLevelColumn = turnoverMax(StockOptionLevel, data=turnover, "Stock Option Level")
turnoverFactors = rbind(turnoverFactors, StockOptionLevelColumn)

TotalWorkingYearsColumn = turnoverMax(TotalWorkingYears, data=turnover, "Total Working Years")
turnoverFactors = rbind(turnoverFactors, TotalWorkingYearsColumn)

TrainingTimesLastYearColumn = turnoverMax(TrainingTimesLastYear, data=turnover, "Training Times Last Year")
turnoverFactors = rbind(turnoverFactors, TrainingTimesLastYearColumn)

WorkLifeBalanceColumn = turnoverMax(WorkLifeBalance, data=turnover, "Work Life Balance")
turnoverFactors = rbind(turnoverFactors, WorkLifeBalanceColumn)

YearsAtCompanyColumn = turnoverMax(YearsAtCompany, data=turnover, "Years At Company")
turnoverFactors = rbind(turnoverFactors, YearsAtCompanyColumn)

YearsInCurrentRoleColumn = turnoverMax(YearsInCurrentRole, data=turnover, "Years In Current Role")
turnoverFactors = rbind(turnoverFactors, YearsInCurrentRoleColumn)

YearsSinceLastPromotionColumn = turnoverMax(YearsSinceLastPromotion, data=turnover, "Years Since Last Promotion")
turnoverFactors = rbind(turnoverFactors, YearsSinceLastPromotionColumn)

YearsWithCurrManagerColumn = turnoverMax(YearsWithCurrManager, data=turnover, "Years With Current Manager")
turnoverFactors = rbind(turnoverFactors, YearsWithCurrManagerColumn)

turnoverFactorsSorted = turnoverFactors  %>% arrange(desc(`Attrition Count`))
turnoverFactorsSorted %>% select(Reason, `Attrition Factor`, `Attrition Count`)

```
## Top 3 Attrition Factors
The top 3 attrition factors are Performance Rating of 3, Stock Option Level of 0, and Business Travel with a factor of Travel_Rarely.  These were obtained by using the group_by function on Attrition and each column name, and doing a count on the number of times Attrition = "Yes". The results are the total number of "Yes" for each factor.
```{r}
turnoverFactorsBarChart = turnoverFactorsSorted %>% filter(Reason == "Performance Rating" | Reason == "Stock Option Level" | Reason == "Business Travel") %>% ggplot(aes(x=reorder(Reason, `Attrition Count`), y=`Attrition Count`, fill=`Attrition Factor`)) + geom_col(position="dodge2") + coord_flip() + xlab("Attrition Reason") + ggtitle("Top 3 Attrition Factors")
ggplotly(turnoverFactorsBarChart)
```

## Reasons by Factor
This shows all Reasons broken down by Factor. We ignore Standard Hours and Over 18 as all employees work standard hours and are over 18.  Use the zoom tool to zoom on columns with multiple factors.
```{r}
turnoverFactorsBarChart = turnoverFactorsSorted %>% ggplot(aes(x=reorder(Reason, `Attrition Count`), y=`Attrition Count`, fill=`Attrition Factor`)) + geom_col(position="dodge2") + coord_flip() + xlab("Attrition Reason") + ggtitle("All Attrition")
ggplotly(turnoverFactorsBarChart, height=1000)
```
##Job Role Specific Trends
```{r}

```

## Predictions of Attrition
I set up a point system to make the data more KNN friendly.  I chose the top factors and assigned them as Personal Points, or Business Points. If an employee has any of these characteristics, they get a point added, else they get 0. More points means an increased chance that Attrition will equal "Yes".

## Personal Points
Age	29
DistanceFromHome	1
Education	3
EducationField	Life Sciences
EnvironmentSatisfaction	1
Gender	Male
JobSatisfaction	
MaritalStatus	Single
NumCompaniesWorked	1
RelationshipSatisfaction	4
TotalWorkingYears	
WorkLifeBalance	3

## Business Points
BusinessTravel	Travel_Rarely	
Department	Research & Development	
JobInvolvement	3	
JobLevel	1	
JobRole	Sales Executive	| Research Scientist
OverTime	Yes	
PerformanceRating	3	
StockOptionLevel	0	
TrainingTimesLastYear	2	
YearsInCurrentRole	0 |	2
YearsSinceLastPromotion	0	
YearsWithCurrManager	0	


```{r}
# Set up the points system
turnoverPoints = turnoverData %>% mutate(
  PersonalPoints = (
    ifelse(Age == 29, 1,0) +
    ifelse(DistanceFromHome == 1, 1,0) +
    ifelse(Education  == 3, 1,0) +
    ifelse(EducationField  == "Life Sciences", 1,0) +
    ifelse(EnvironmentSatisfaction  == 1, 1,0) +
    ifelse(Gender  == "Male", 1,0) +
    ifelse(MaritalStatus  == "Single", 1,0) +
    ifelse(NumCompaniesWorked  == 1, 1,0) +
    ifelse(RelationshipSatisfaction  == 4, 1,0) +
    ifelse(WorkLifeBalance  == 3, 1,0)
  ),
  BusinessPoints = (
    ifelse(Department == "Research % Development", 1,0) +
    ifelse(JobInvolvement == 3, 1,0) +
    ifelse(JobLevel == 1, 1,0) +
    ifelse(JobRole == "Sales Executive" | JobRole == "Research Scientist", 1,0) +
    ifelse(OverTime == "Yes", 1, 0) +
    ifelse(PerformanceRating == 3, 1,0) +
    ifelse(StockOptionLevel == 0, 1,0) +
    ifelse(TrainingTimesLastYear == 2, 1,0) +
    ifelse(YearsInCurrentRole == 0 | YearsInCurrentRole == 2, 1,0) +
    ifelse(YearsSinceLastPromotion == 0, 1,0) +
    ifelse(YearsWithCurrManager == 0, 1,0)
  )
)

turnoverPoints = turnoverPoints %>% select(ID, PersonalPoints, BusinessPoints, Attrition)
turnoverPoints %>% group_by(Attrition, PersonalPoints, BusinessPoints) %>% summarize(count = n())

```
## KNN to Predict Attrition
```{r}
attritionTrain = data.frame(
  as.factor(turnoverPoints$ID),
  as.factor(turnoverPoints$PersonalPoints),
  as.factor(turnoverPoints$BusinessPoints),
  as.factor(turnoverPoints$Attrition)
)
names(attritionTrain) = c("ID", "Personal", "Business", "Attrition")



# Split data into test/train and build KNN
set.seed(10)
iterations = 100
splitPerc = .70
trainIndices = sample(1:dim(attritionTrain)[1],round(splitPerc * dim(attritionTrain)[1]))
train = attritionTrain[trainIndices,]
test = attritionTrain[-trainIndices,]
accs = data.frame(accuracy = numeric(iterations), k = numeric(iterations))
for(i in 1:iterations)
{
    classifications = knn(train[,c(2,3)],test[,c(2,3)],train$Attrition, prob = TRUE, k = i)
    table(test$Attrition,classifications)
    CM = confusionMatrix(table(test$Attrition,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}

#plot used to review best K values
plot(accs$k,accs$accuracy, type = "l", xlab = "k", ylab="Accuracy")

classifications = knn(train[,c(2,3)],test[,c(2,3)],train$Attrition, prob = TRUE, k = accs$k[which.max(accs$accuracy)])
CM = confusionMatrix(table(test$Attrition,classifications))
CM

# Open csv file for our predictions
attritionTest = read.csv('CaseStudy2CompSet No Attrition.csv')
# Set up the points system
turnoverPoints = attritionTest %>% mutate(
  PersonalPoints = (
    ifelse(Age == 29, 1,0) +
    ifelse(DistanceFromHome == 1, 1,0) +
    ifelse(Education  == 3, 1,0) +
    ifelse(EducationField  == "Life Sciences", 1,0) +
    ifelse(EnvironmentSatisfaction  == 1, 1,0) +
    ifelse(Gender  == "Male", 1,0) +
    ifelse(MaritalStatus  == "Single", 1,0) +
    ifelse(NumCompaniesWorked  == 1, 1,0) +
    ifelse(RelationshipSatisfaction  == 4, 1,0) +
    ifelse(WorkLifeBalance  == 3, 1,0)
  ),
  BusinessPoints = (
    ifelse(Department == "Research % Development", 1,0) +
    ifelse(JobInvolvement == 3, 1,0) +
    ifelse(JobLevel == 1, 1,0) +
    ifelse(JobRole == "Sales Executive" | JobRole == "Research Scientist", 1,0) +
    ifelse(OverTime == "Yes", 1, 0) +
    ifelse(PerformanceRating == 3, 1,0) +
    ifelse(StockOptionLevel == 0, 1,0) +
    ifelse(TrainingTimesLastYear == 2, 1,0) +
    ifelse(YearsInCurrentRole == 0 | YearsInCurrentRole == 2, 1,0) +
    ifelse(YearsSinceLastPromotion == 0, 1,0) +
    ifelse(YearsWithCurrManager == 0, 1,0)
  )
)

attritionTest = data.frame(
  as.factor(turnoverPoints$ID),
  as.factor(turnoverPoints$PersonalPoints),
  as.factor(turnoverPoints$BusinessPoints)
)
names(attritionTest) = c("ID", "Personal", "Business")

#
for (i in attritionTest[1]) {
  df1 = attritionTest %>% filter(ID == i) %>% select(Personal, Business)
  df1knn = knn(attritionTrain[,c(2,3)], df1, attritionTrain$Attrition, k = 10, prob = TRUE)
}

finalPredictions = data.frame(attritionTest$ID[1:300], prediction=df1knn[1:300])
names(finalPredictions) = c("ID", "Prediction")
finalPredictions %>% filter(Prediction == "Yes")
finalPredictions
write_excel_csv(finalPredictions, "Case2PredictionsMauldin Attrition.csv")

```

## Predictions of Salary
Here, I am using cross validation to predict the salary of a test dataset.
```{r}
testSalary = read.csv('CaseStudy2CompSet No Salary.csv')

trainSalary = data.frame(turnoverData[33], turnoverData[16], turnoverData[20])
trainSalary

set.seed(4)
trainObs = sample(seq(1,dim(trainSalary)[1]),round(.75*dim(trainSalary)[1]),replace = FALSE)
train = trainSalary[trainObs,]
train
test = testSalary
test
fit1 = lm(MonthlyIncome~YearsAtCompany+JobLevel, data = trainSalary)
summary(fit1)
fit1Preds = predict(fit1, newdata = test)
finalSalaryPredictions = data.frame(ID=test$ID, Prediction=fit1Preds)
finalSalaryPredictions
write_excel_csv(finalSalaryPredictions, "Case2PredictionsMauldin Salary.csv")

```